Towards Patient-Specific Neural Field Modeling using Micro-Electrode Data

### 2011-05-07 ltx/data_paper/DataPaper.tex
changed the kernel estimate plot
added another plot for kernel support estimation
minor edit

\usepackage{amsmath,amssymb,amsfonts,epsfig} % Typical maths resource packages

%\newif\ifpdf
%\ifx\pdfoutput\undefined
%\pdffalse % we are not running PDFLaTeX
%\else
%\pdfoutput=1 % we are running PDFLaTeX
%\pdftrue
%\fi

% \ifpdf
% \usepackage[pdftex]{graphicx}
% \else
% \usepackage{graphicx}
% \fi
\usepackage{color}
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}

%\newcommand{\red}{\textcolor{red}}
%\newcommand{\blue}{\textcolor{blue}}
\newcommand{\cyan}{\textcolor{cyan}}

@@

The neural field equations used as the parametric form of the model are based on the influential papers from Wilson and Cowan~\cite{Wilson1973}, and Amari~\cite{Amari1977}. This class of model describes a continuous cortical sheet or surface, relating mean firing rates of pre and post-synaptic neural populations. The neural field, $v\left( {\mathbf{r},t} \right)$, at position $\mathbf{r}$ and time $t$, is the aggregated post-synaptic potentials described by the \dean{(temporal convolution - I'd rather say integral equation	)} integral equation 

@@

$\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function \parham{(you can keep it here if you wish as it doesn't change the integral equation because of the upper limit. I would remove it)}. $g(\cdot)$ describes the mean firing rate. In practice, the firing rate can rarely be considered as a purely determinstic process because \dean{(can someone write down why?)}. We thus choose to introduce a spatially coloured space-time Wiener process $W(\mathbf{r},t)$ to model the flactuations in the mean firing rate and decompose $g(\cdot)$ as

@@

where $\sigma_W \ge 0$ is a measure of the introduced space-time randomness. For $\sigma_W = 0$ a pure deterministic process is recovered. As a result of the random component, for $\sigma_W > 0$, the membrane voltage $v(\mathbf{r},t)$ is itself a stochastic process. Assuming an infinite propagation velocity for action potentials within the field, the deterministic incoming firing rate is described by a spatial convolution \dean{citation?}

@@

 \textrm{d}v(\mathbf{r},t) + \zeta v(\mathbf{r},t)\textrm{d}t = \tilde{g}(\mathbf{r},t)\textrm{d}t + \textrm{d}W(\mathbf{r},t)  ~~~~ t \ge 0, v(\mathbf{r},0) = \cyan{GP?}

@@

	\boldsymbol\Sigma_e \triangleq \sigma_d^2\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 

@@

Note that $\boldsymbol\Sigma_e=\sigma_d^2\tilde{\boldsymbol\Sigma}_e$ (\dean{$\sigma_W$ notation from 18}\parham{I fixed it by defining $\sigma_d^2 =T_s\sigma^2_W$ in the model derivation}) where $\tilde{\boldsymbol\Sigma}_e$ is constant and can be calculated beforehand as the support of the disturbance covariance function is assumed to be known. By expanding the exponent and taking $2\ln$ we get

@@

noting that variable $\boldsymbol\Xi_4$ is symmetric by construction, equating \eqref{eq:Qpartialtheta} to the zero vector yields

@@

We comment on the condition under which $\boldsymbol\Xi_4$ is invertible later on. The second derivative is negative definite at this stationary point, and hence it is a global maximizer of $\mathcal{Q}$-function. Now maximizing with respect to $\sigma_{\epsilon}^2$ we get

@@

meaning \eqref{eq:SecondDerivativexi} is negative and $\hat{\xi}$ represents a maximum. Equation \eqref{eq:SecondDerivativetheta} is negative where $\boldsymbol\Xi_4$ is positive definite requiring $\mathbf q(\mathbf x_t)$ to be of  rank $n_{\theta}$ \cite{Golub1996}. The matrix $\boldsymbol\Psi(\mathbf r')$ defined by \eqref{eq:DefPsi} is of rank $n_{\theta}$ as each row of the matrix is the set of equally spaced Gaussian basis functions with equal widths, consequently each row can be considered as a Riesz basis function for the subspaces they generate \cite{Ferrari2004}. (note that multiplying by $\mathbf \Gamma^{-1}$ doesnt change the rank). However, $q(\mathbf x_t)$ defined in \eqref{eq:QmatrixForSigmapoints} cannot be computed analytically as the inner product of a Gaussian and a sigmoid cannot be solved analytically making it difficult to calculate the rank of $q(\mathbf x_t)$ . Therefore, we recommend to check the rank of $q(\mathbf x_t)$ at each iteration.

@@

From \eqref{eq:SecondDerivativenoise} and \eqref{eq:SecondDerivativedisturbance}, observation noise and disturbance covariance estimates clearly represent maxima of the $\mathcal{Q}$-function.

@@

\begin{itemize}
	\item comment on how the e step is UKF and M-step is kinda like the EKF, and we dont use EKF in E-step since UKF gives better estimates and the linearisation in M-step would be more accurate. 
	\item comment on EM has been used with EKS by Ghahramani which has poorer convergance property (I guess) using EM with URTSS
\end{itemize}

>>>>

\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages

\newif\ifpdf
\ifx\pdfoutput\undefined
\pdffalse % we are not running PDFLaTeX
\else
\pdfoutput=1 % we are running PDFLaTeX
\pdftrue
\fi

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi
\usepackage{color}
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\red}{\textcolor{red}}
\newcommand{\blue}{\textcolor{blue}}
\newcommand{\cyan}{\textcolor{cyan}}

@@

The neural field equations used as the parametric form of the model are based on the influential papers from Wilson and Cowan~\cite{Wilson1973}, and Amari~\cite{Amari1977}. This class of model describes a continuous cortical sheet or surface, relating mean firing rates of pre and post-synaptic neural populations. The neural field, $v\left( {\mathbf{r},t} \right)$, at position $\mathbf{r}$ and time $t$, is the aggregated post-synaptic potentials described by the \red{(temporal convolution - I'd rather say integral equation	)} integral equation 

@@

$\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function \blue{(you can keep it here if you wish as it doesn't change the integral equation because of the upper limit. I would remove it)}. $g(\cdot)$ describes the mean firing rate. In practice, the firing rate can rarely be considered as a purely determinstic process because \red{(can someone write down why?)}. We thus choose to introduce a spatially coloured space-time Wiener process $W(\mathbf{r},t)$ to model the flactuations in the mean firing rate and decompose $g(\cdot)$ as

@@

where $\sigma_W \ge 0$ is a measure of the introduced space-time randomness. For $\sigma_W = 0$ a pure deterministic process is recovered. As a result of the random component, for $\sigma_W > 0$, the membrane voltage $v(\mathbf{r},t)$ is itself a stochastic process. Assuming an infinite propagation velocity for action potentials within the field, the deterministic incoming firing rate is described by a spatial convolution \red{citation?}

@@

 \textrm{d}v(\mathbf{r},t) + \zeta v(\mathbf{r},t)\textrm{d}t = \tilde{g}(\mathbf{r},t)\textrm{d}t + \textrm{d}W(\mathbf{r},t)  ~~~~ t \ge 0, v(\mathbf{r},0) = \red{GP?}

@@

	\boldsymbol\Sigma_e \triangleq T_s\sigma_W^2\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 

@@

Note that $\boldsymbol\Sigma_e=\sigma_d^2\tilde{\boldsymbol\Sigma}_e$ (\dean{$\sigma_W$ notation from 18}) where $\tilde{\boldsymbol\Sigma}_e$ is constant and can be calculated beforehand as the support of the disturbance covariance function is assumed to be known. By expanding the exponent and taking $2\ln$ we get

@@

variable $\boldsymbol\Xi_4$ is symmetric and positive-definite by construction, equating \eqref{eq:Qpartialtheta} to the zero vector yields

@@

The second derivative is negative definite at this stationary point, and hence it is a global maximizer of $\mathcal{Q}$-function. Now maximizing with respect to $\sigma_{\epsilon}^2$ we get

@@

meaning \eqref{eq:SecondDerivativexi} is negative and $\hat{\xi}$ represents a maximum. Equation \eqref{eq:SecondDerivativetheta} is negative where $\boldsymbol\Xi_4$ is positive definite requiring $\mathbf q(\mathbf x_t)$ to be of  rank $n_{\theta}$ \cite{Golub1996}, 

@@

\blue{don't know how to prove $q$ is of rank $n_{\theta}$} According to \eqref{eq:SecondDerivativenoise} and \eqref{eq:SecondDerivativedisturbance}, observation noise and disturbance covariance estimates clearly represent maxima of the $\mathcal{Q}$-function.

@@

\begin{itemize}
	\item comment on how the e step is UKF and M-step is kinda like the EKF, and we dont use EKF in E-step since UKF gives better estimates and the linearisation in M-step would be more accurate.  
\end{itemize}
